---
title: "Stochastic Volatility Modeling"
author: "Sewon Chang"
date: "2024-08-28"
output: github_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```
```{r}
library(quantmod)
library(ghyp)
```



Start with extracting the daily exchange-rate series for the currencies 
EUR, RUB, and CNY from Yahoo from 2017 to 2022 using the getSymbols function
from quantmod

```{r}
getSymbols("EUR=X",src="yahoo",from="2017-01-01", to="2022-01-01")
EUR <- `EUR=X`
getSymbols("RUB=X",src="yahoo",from="2017-01-01", to="2022-01-01")
RUB <- `RUB=X`
getSymbols("CNY=X",src="yahoo",from="2017-01-01", to="2022-01-01")
CNY <- `CNY=X`
```



We compute daily log returns using adjusted price of the exchange rates, 
and omit any empty returns

```{r}
all_returns <- cbind(diff(log(EUR$`EUR=X.Adjusted`)), diff(log(RUB$`RUB=X.Adjusted`)), ... = diff(log(CNY$`CNY=X.Adjusted`)))
all_Returns <- na.omit(all_returns)

EUR_Returns <- all_Returns$EUR.X.Adjusted
RUB_Returns <- all_Returns$RUB.X.Adjusted
CNY_Returns <- all_Returns$CNY.X.Adjusted

```



Set the five-year period to be from Jan 1st, 2017 to Dec 31st, 2021.
The training set is then from Jan 1st, 2017 to Dec 30th, 2020 and
the test set is from Jan 1st, 2020 to Dec 31st, 2021.

```{r}
start_Tr <- as.Date("2017-01-01")
end_Tr <- as.Date("2020-12-31")

returnsDF <- data.frame(EUR_Returns, CNY_Returns, RUB_Returns)
returnsDF$Date <- as.Date(rownames(returnsDF))

Training <- subset(returnsDF, Date >= start_Tr, Date <= end_Tr)
Training <- subset(Training, select = -c(Date))
head(Training)

start_test <- as.Date("2021-01-01")
end_test <- as.Date("2021-12-31")

Test <- subset(returnsDF, Date >= start_test, Date <= end_test)
Test <- subset(Test, select = -c(Date))
```



Apply generalized hyperbolic distribution, normal inverse Gaussian,
skewed student-t, and variance gamma distribution, respectively on the 
multivariate data sets using fitting functions from ghyp package

Then, the sum of the densities are calculated

```{r}
Training <- sapply(Training, as.numeric)
Test <- sapply(Test, as.numeric)
```
```{r, results='hide'}
GH_Tr <- fit.ghypmv(Training)
NIG_Tr <- fit.NIGmv(Training)
t_Tr <- fit.tmv(Training)
VG_Tr <- fit.VGmv(Training)

GH_Test <- fit.ghypmv(Test)
NIG_Test <- fit.NIGmv(Test)
t_Test <- fit.tmv(Test)
VG_Test <- fit.VGmv(Test)

GH_Test_LL <- sum(dghyp(Test, GH_Tr, logvalue = TRUE))
NIG_Test_LL <- sum(dghyp(Test, NIG_Tr, logvalue = TRUE))
t_Test_LL <- sum(dghyp(Test, t_Tr, logvalue = TRUE))
VG_Test_LL <- sum(dghyp(Test, VG_Tr, logvalue = TRUE))
```


Log-likelihood of Training and Test Sets are shown in the table below:

```{r}


a <- logLik(GH_Tr)
b <- logLik(NIG_Tr)
c <- logLik(t_Tr)
d <- logLik(VG_Tr)

e <- GH_Test_LL
f <- NIG_Test_LL
g <- t_Test_LL
h <- VG_Test_LL

Table <- matrix(c(a,b,c,d,e,f,g,h), ncol=2)
colnames(Table) <- c("Training", "Test")
rownames(Table) <- c("ghyp", "NIG", "t", "VG")

Table
```

Volatilies of each GH distribution from the diagonals of the covariance matrix

```{r}
GH_vol <- sqrt(diag(GH_Tr@sigma))
NIG_vol <- sqrt(diag(NIG_Tr@sigma))
t_vol <- sqrt(diag(t_Tr@sigma))
VG_vol <- sqrt(diag(VG_Tr@sigma))

vols <- data.frame(GH_vol, NIG_vol, t_vol, VG_vol)
vols
```
Note that student-t is the most volatile, followed closely by GH, then normal inverse, with 
VG being the least volatile.


Use function coef to find parameters of each distributions.

```{r}
d <- 3

GH_co <- coef(GH_Tr, type = "chi.psi")

lambda_GH <- GH_co$lambda
chi_GH <- GH_co$chi
two_GH <- numeric()
psi_GH <- GH_co$psi

mu_GH <- GH_co$mu
gamma_GH <- GH_co$gamma
sigma_GH <- GH_co$sigma




NIG_co <- coef(NIG_Tr, type = "chi.psi")

lambda_NIG <- NIG_co$lambda
chi_NIG <- NIG_co$chi
two_NIG <- numeric()
psi_NIG <- NIG_co$psi

mu_NIG <- NIG_co$mu
gamma_NIG <- NIG_co$gamma
sigma_NIG <- NIG_co$sigma





t_co <- coef(t_Tr, type = "chi.psi")

lambda_t <- t_co$lambda
chi_t <- t_co$chi
two_t <- numeric()
psi_t <- t_co$psi

mu_t <- t_co$mu
gamma_t <- t_co$gamma
sigma_t <- t_co$sigma




VG_co <- coef(VG_Tr, type = "chi.psi")

lambda_VG <- VG_co$lambda
chi_VG <- VG_co$chi
two_VG <- numeric()
psi_VG <- VG_co$psi

mu_VG <- VG_co$mu
gamma_VG <- VG_co$gamma
sigma_VG <- VG_co$sigma


one_GH <- lambda_GH - d/2
one_NIG <- lambda_NIG - d/2
one_t <- lambda_t - d/2
one_VG <- lambda_VG - d/2
```


Using the obtained parameters, calculate the mixing variable W given
by a Generalized Inverse Gaussian distribution (GIG) as they are influenced
by the log returns

```{r}
for(i in 1:1296){
two_GH[i] <- (t(Training[i,]-mu_GH) %*% solve(sigma_GH) %*% (Training[i,]-mu_GH)) + chi_GH
two_NIG[i] <- (t(Training[i,]-mu_NIG) %*% solve(sigma_NIG) %*% (Training[i,]-mu_NIG)) + chi_NIG
two_t[i] <- (t(Training[i,]-mu_t) %*% solve(sigma_t) %*% (Training[i,]-mu_t)) + chi_t
two_VG[i] <- (t(Training[i,]-mu_VG) %*% solve(sigma_VG) %*% (Training[i,]-mu_VG)) + chi_VG
}

three_GH <- psi_GH + t(gamma_GH) %*% solve(sigma_GH) %*% gamma_GH
three_NIG <- psi_NIG + t(gamma_NIG) %*% solve(sigma_NIG) %*% gamma_NIG
three_t <- psi_t + t(gamma_t) %*% solve(sigma_t) %*% gamma_t
three_VG <- psi_VG + t(gamma_VG) %*% solve(sigma_VG) %*% gamma_VG


GH_Egig <- Egig(one_GH, two_GH, three_GH)
NIG_Egig <- Egig(one_NIG, two_NIG, three_NIG)
t_Egig <- Egig(one_t, two_t, three_t)
VG_Egig <- Egig(one_VG, two_VG, three_VG)


Date <- returnsDF$Date
GH <- data.frame(Date, GH_Egig)
NIG <- data.frame(Date, NIG_Egig)
t <- data.frame(Date, t_Egig)
VG <- data.frame(Date, VG_Egig)
```


Plot time series graphs of the resulting mixing GIG distribution
to obtain stochastic volatility models for each tested distribution

```{r}
plot(GH, type = "line", col = "turquoise", xlab = "Year", ylab = "Volatility", lwd = 3)
lines(NIG, col = "red", lwd = 2)
lines(t, col = "orange")
lines(VG, col = "blue")
legend("topleft", legend=c("GH", "normal inverse",
                            "skewed student-t", "variance gamma"), fill=c("turquoise",
                            "red", "orange","blue"))
```
Notice that the volatilities reach their max on 2020-03-09
With VG having the lowest volatility of 5.733866, NIG being 8.342398,
GH being 16.68439 and student-t being the highest with 16.87061



Below is also a smoothed version of the graph:

```{r}
GH_r <- data.frame(Date, rollmean(GH$GH_Egig, 50, na.pad = TRUE))
NIG_r <- data.frame(Date, rollmean(NIG$NIG_Egig, 50, na.pad = TRUE))
t_r <- data.frame(Date, rollmean(t$t_Egig, 50, na.pad = TRUE))
VG_r <- data.frame(Date, rollmean(VG$VG_Egig, 50, na.pad = TRUE))

plot(GH_r, type = "line", col = "turquoise", xlab = "Year", ylab = "Volatility", main = "Stochastic Volatility model of exchange-rate log-returns data",lwd = 2)
lines(NIG_r, col = "red")
lines(t_r, col = "orange")
lines(VG_r, col = "blue")
legend("topleft", legend=c("GH", "normal inverse",
                            "skewed student-t", "variance gamma"), fill=c("turquoise",
                            "red", "orange","blue"))
```
